---
title: "Assignment 4"
author: "Elmera Azadpour"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(chron)
library(kableExtra)
library(effsize)
library(ggpubr)
library(broom)
```


```{r, error=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
hares <- read_csv(here("data","bonanza_hares.csv"))
```

### Introduction:
This mini exploratory report is on juvenile snowshoe hares in Bonanza Creek Experimental Forest (Long Term Ecological Research site). The Bonanza Creek LTER is a member of the U.S. LTER Network which is supported by the National Science Foundation and by the USDA Forest Service, Pacific Northwest Research Station. The Bonanza Creek Experimental Forest (BCEF) was established in 1963 is a 50 km^2^ experimental forest located approximately 20 km southwest of Fairbanks, Alaska and located on land owned by the State of Alaska. The Forest is within the Tanana Valley State Forest, a unit managed by the Division of Forestry, State of Alaska. This report covers analyses including: annual juvenile hare trap counts, comparison of juvenile hare weights by sex and site, juvenile weight comparison between male & female snowshoe hares, and an analysis of the relationship between juvenile snowshoe hare hind foot length. (information regarding teh Bonaza Creek LTER network was provided from https://www.lter.uaf.edu/research/study-sites-bcef)

### Data and analyses:
This analysis contains data with counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012. This report analyses: differences in annual juvenile hare trap counts (Fig. 1), comparison of juvenile hare weights by sex and site (Fig 2), comparison of mean weights for juvenile male and female snowshoe hares (Table 1), comparison of the relationship between juvenile weight & hind foot length (Fig. 3), and a linear regression describing how juvenile snowshoe hare weight changes with respect to hind foot length (Fig. 4). Along with the linear model, tests for normality, two-sided test, Pearson's *r*, and R^2^ analyses were also conducted. 

## 1) Exploratory findings
### Annual juvenile hare trap counts:
```{r, fig.cap = "Figure 1. Annual juvenile hare trap counts (1999 - 2012)", error=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
## break up date using lubridate

hares_juvinile <- hares %>% 
  mutate(Date = mdy(date))%>% 
  mutate(Year = year(Date)) %>% 
  filter(age == "j") %>% 
  group_by(Year) %>% 
  count()

ggplot(data = hares_juvinile,
       aes(x=Year, y= n)) + 
  geom_line(col= "red")  +
  geom_point(col = "red") +
  theme_minimal() +
  labs(x= "Year", 
       y='Count of hare trappings', 
       title = "Annual juvenile hare trap counts (1999 - 2012)") +
  scale_x_continuous(breaks=seq(1999, 2012, 2)) 

# mean(hares_juvinile$n)
# median(hares_juvinile$n)
```
#### Major takeaways from Figure 1: 
The max count of juvenile hare trappings was 126 in 1999. The min count of juvenile hare trappings was 2 in 2010. The mean annual number of juvenile hares trapped was 31.5. The median annual number of juvenile hares trapped is 18.5. More generally, in more recent years (2005-2012), hare trapping have fluctuated quite a bit showing sharp increases followed by decreases. 

## 2) Visualization of juvenile hare weights.
```{r, fig.cap = "Figure 2. Mean juvenile hare weights by sex and site (1999 - 2012)", error=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
hares_weight <- hares %>% 
  filter(age == "j") %>% 
  select(c("grid", "sex", "weight")) 

hares_weight$grid <- as.factor(hares_weight$grid)
grid.labs <- c("Bonbs", "Bonbs", "Bonrip")
names(grid.labs) <- c("bonbs", "bonmat", "bonrip")

ggplot(hares_weight, 
       aes(x=sex, y= weight)) +
  geom_boxplot(aes(fill=sex)) +
  facet_wrap(~grid, labeller = labeller(grid = grid.labs)) + 
  theme_minimal() +
  labs(x= "Sex", 
       y='Weight (g)', 
       title = "Juvenile hare weights by sex and site (1999 - 2012)") +
  scale_x_discrete(labels= c("Female", "Male", "NA")) + 
  scale_y_continuous(breaks=seq(0, 1700, 300)) +
  theme(legend.position = "none")

```
#### Major takeaways from Figure 2
This graph depictsjuvenile hare weight by sex across the three sites. Site bonbs has the max  weight for both male and female hares. In contrast, site bonrip has the lowest  weight of female and male hares. Additinally, this graph depicts the upper and lower quartiles of each sex at the differnt sites. This graph also includes NA values. 

## 3) Juvenile weight comparison of male & female snowshoe hares
```{r, error=FALSE, warning=FALSE, message=FALSE, echo = FALSE}
hares_weight_mf <- hares %>% 
  filter(age == "j") %>% 
  select(c("grid", "sex", "weight")) %>% 
  group_by(sex, grid) %>% 
  summarise(mean_weight = mean(weight, na.rm=T), 
            median_weight = median(weight, na.rm=T),
            sd_weight = sd(weight, na.rm=T),
            n = n())

hares_weight_mf %>%  
  kable(col.names = c("Sex","Site","Mean weight (g)",
                      "Median weight (g)",
                      "Standard deviation",
                      "Sample size (n)"), 
        caption = "Table 1: summary statistics of juvenile male and female snowshoe hare 1999 - 2012") %>% 
  kable_styling(full_width = FALSE) 


```

```{r,include = FALSE}
# Histograms
ggplot(data = hares_weight_mf, aes(x = mean_weight)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex, scales = "free")

# QQ Plots
ggplot(data= hares_weight_mf, aes(sample = mean_weight)) +
  geom_qq() +
  facet_wrap(~sex)

# two sample t-test
hares_weight_f <- hares_weight_mf %>% 
  select(c("sex", "grid", "mean_weight")) %>% 
  filter(sex == "f") %>% 
  pull(mean_weight)

hares_weight_m <- hares_weight_mf %>% 
  select(c("sex", "grid", "mean_weight")) %>% 
  filter(sex == "m") %>% 
  pull(mean_weight)

t.test(hares_weight_f, hares_weight_m)


# cohens d/ effect size
cohen.d(hares_weight_m, hares_weight_f)
```
#### Major takeaway from means comparison output:
The actual difference in means from juvenile male and female snowshoe hares is 99.3572. The outcomes of the two sample t-test indicates there is a pretty decent chance  (30%) of randomly selecting two samples from populations with the same mean that are this different by chance. The cohend d (effect size) is 0.9879, a large effect size. 



## 4) Examining relationship between juvenile weight & hind foot length.
```{r, fig.cap = "Figure 3. Juvenile hare weights vs hind foot length from 1999 - 2012", error=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
hares_hind <- hares %>% 
  filter(age == "j") %>% 
  select(c("weight", "hindft"))


ggplot(hares_hind, 
       aes(x=weight, y=hindft)) + 
  geom_point() +
  theme_minimal() +
  labs(x= "Weight (g)", 
       y='Hind foot length (mm)', 
       title = "Juvenile hare weight vs. hind foot length (1999 - 2012)") + 
  scale_x_continuous(breaks=seq(0, 1800, 400)) 

```

## 5) Linear model output
```{r, include = FALSE}
## linear regression 
hares_lm <- lm(weight ~ hindft, data = hares_hind)

# Return the complete overview:
summary(hares_lm)

# We can use the broom::tidy() function to get the model outputs in nice data frame format:
hares_lm_tidy <- broom::tidy(hares_lm)

# Get the intercept: 
hares_int <- hares_lm_tidy$estimate[1]
hares_int

# Then to get the flipper_length coefficient:
hares_coef <- hares_lm_tidy$estimate[2]
hares_coef


#What about getting some other model information (degrees of freedom, F-statistic, p-value, etc.)?
#Many of these statistical outcomes can be accessed more easily using broom::glance().
# Metrics at a glance: 
hares_lm_out <- broom::glance(hares_lm)
hares_lm_out


# Explore model assumptions
plot(hares_lm)

# Pearson’s r
hares_cor <- cor.test(hares_hind$weight, hares_hind$hindft)
hares_cor
```


```{r, fig.cap = "Figure 4. Linear model of weight (g) vs hind foot length  (mm) in juvenile snowshoe hare (1999 - 2012)", error=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
# Visualize the model
ggplot(data = hares_hind, aes(x = weight, y = hindft)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 230, label.y = 160)
```

#### Outcomes of linear regression:
Simple linear regression was used to explore between juvenile snowshoe hare weight (g) and hare hind foot length (mm) across all  sites, and including both male and female snowshoe hares. A significant regression model was found ($\beta$ = `r round(hares_coef,3)`, F(`r hares_lm_out$df`,`r hares_lm_out$df.residual`) = `r round(hares_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(hares_lm_out$r.squared,3)`. The slope of the linear model is 9.523 and the y-intercept is -279.3. This slope of the linear model suggests that on average, we expect that weight to increase by 9.523 grams for each 1 mm of increase in hind foot length. Additionally, when conducting a linear model, there are assumptions such as, linearly related variables, normally distributed, Homoscedasticity and iid residuals (no serial correlation). When looking at the Residual Variances, for Homoscedasticity, it appears residuals do not appear randomly distributed evenly around the horizontal dotted line. Additionally, in the QQ plot, the residuals do not appear to be entirely normally distributed -- theres quite a few points on the tails that drift from the dotted line. This is probably partly why we see a more moderate R value (0.547). 


#### Summary of juvenile hares exploration:
+ The max count of juvenile hare trappings was 126 in 1999. The min count of juvenile hare trappings was 2 in 2010. The mean annual number of juvenile hares trapped was 31.5. The median annual number of juvenile hares trapped is 18.5. More generally, in more recent years (2005-2012), hare trapping have fluctuated quite a bit showing sharp increases followed by decreases.
+ The actual difference in means from juvenile male and female snowshoe hares is 99.3572. The outcomes of the two sample t-test indicates there is a pretty decent chance  (30%) of randomly selecting two samples from populations with the same mean that are this different by chance. The cohend d (effect size) is 0.9879, a large effect size. 
+ This slope of the linear model suggests that on average, we expect that weight to increase by 9.523 grams for each 1 mm of increase in juvenile hind foot length.
+ Obtained a moderate persons *r* =  0.547 value between weights and hind foot length


#### Citations:
+ Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

+ Link to metadata: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-bnz.55.22
